### 2017-07-21 case-study/change_frequency.tex
DRY up

% \begin{tikzpicture}
%   \begin{axis}[
%     title=\dataSourceName\ API Changes,
%     date coordinates in=x,
%     stack plots=y,
%     area style,
%     xtick = data,
%     xlabel={Release Dates},
%     ylabel={No. of Changes},
%     xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%     enlarge x limits=false,
%     height=0.25\textheight,
%     width=0.9\linewidth,
%     legend style={legend pos=north west,font=\tiny, legend columns=2}
%     ]
    \addplot[fill=green] table[x=Date, y=AddAPIElements]{plotdata/\dataSourceFileName}
    \closedcycle;
    \addplot[fill=red] table[x=Date, y=RemoveAPIElements]{plotdata/\dataSourceFileName}
    \closedcycle;
    \addplot[fill=orange] table[x=Date, y=ChangeAPIElements]{plotdata/\dataSourceFileName}
    \closedcycle;
    \addplot[fill=gray] table[x=Date, y=ChangeFieldDataType]{plotdata/\dataSourceFileName}
    \closedcycle;
    \addplot[fill=violet] table[x=Date, y=ErrorConditionChange]{plotdata/\dataSourceFileName}
    \closedcycle;
    \addplot[fill=yellow] table[x=Date, y=ChangeAuthentication]{plotdata/\dataSourceFileName}
    \closedcycle;
    \addplot[fill=magenta] table[x=Date, y=BehaviorChange]{plotdata/\dataSourceFileName}
    \closedcycle;
    \addplot[fill=blue] table[x=Date, y=ChangeAuthorization]{plotdata/\dataSourceFileName}
    \closedcycle;

    \legend{AddAPIElements, RemoveAPIElements, ChangeAPIElements,
        ChangeFieldDataType, ErrorConditionChange,
        ChangeAuthentication, BehaviorChange, ChangeAuthorization}

%   \end{axis}
% \end{tikzpicture}