### 2017-07-22 case_study/change_frequency.tex
adds the case study paper

\addplot[fill=green] table[x=Date, y=AddAPIElements]{plotdata/\dataSourceFileName}
\closedcycle;
\addplot[fill=red] table[x=Date, y=RemoveAPIElements]{plotdata/\dataSourceFileName}
\closedcycle;
\addplot[fill=orange] table[x=Date, y=ChangeAPIElements]{plotdata/\dataSourceFileName}
\closedcycle;
\addplot[fill=gray] table[x=Date, y=ChangeFieldDataType]{plotdata/\dataSourceFileName}
\closedcycle;
\addplot[fill=violet] table[x=Date, y=ErrorConditionChange]{plotdata/\dataSourceFileName}
\closedcycle;
\addplot[fill=yellow] table[x=Date, y=ChangeAuthentication]{plotdata/\dataSourceFileName}
\closedcycle;
\addplot[fill=magenta] table[x=Date, y=BehaviorChange]{plotdata/\dataSourceFileName}
\closedcycle;
\addplot[fill=blue] table[x=Date, y=ChangeAuthorization]{plotdata/\dataSourceFileName}
\closedcycle;

\legend{AddAPIElements, RemoveAPIElements, ChangeAPIElements,
  ChangeFieldDataType, ErrorConditionChange,
  ChangeAuthentication, BehaviorChange, ChangeAuthorization}