### 2017-07-24 appendix.tex
almost done with conclusion

\chapter{Source Code Fragments of the Spy from Chapter~\ref{chapter:demo_paper}}

\lstdefinelanguage{JavaScript}{
  keywords={typeof, new, true, false, catch, function, return, null, catch, switch, var, if, in, while, do, else, case, break},
  keywordstyle=\color{blue}\bfseries,
  ndkeywords={class, export, boolean, throw, implements, import, this},
  ndkeywordstyle=\color{darkgray}\bfseries,
  identifierstyle=\color{black},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{purple}\ttfamily,
  stringstyle=\color{red}\ttfamily,
  morestring=[b]',
  morestring=[b]"
}

\lstset{
   language=JavaScript,
   %backgroundcolor=\color{lightgray},
   extendedchars=true,
   basicstyle=\footnotesize\ttfamily,
   showstringspaces=false,
   showspaces=false,
   numbers=left,
   numberstyle=\footnotesize,
   numbersep=9pt,
   tabsize=2,
   breaklines=true,
   showtabs=false,
   captionpos=b
}

\begin{lstlisting}[caption=index.js]
(function() {
  var ApiThough;

  ApiThough = require('./lib/api_through');

  process.env.NODE_TLS_REJECT_UNAUTHORIZED = "0";

  process.on('uncaughtException', function(err) {
    console.error('uncaughtException');
    console.error(err);
    return console.error(err.stack);
  });

  new ApiThough().start();

}).call(this);
\end{lstlisting}

\begin{lstlisting}[caption=lib/api\_through.js]
(function() {
  var ApiThrough, PassThrough;

  module.exports = function() {
    return new ApiThrough();
  };

  PassThrough = require('stream').PassThrough;

  ApiThrough = (function() {
    var Proxy, SSLCert, fs;

    Proxy = require('http-mitm-proxy');

    SSLCert = require('./ssl_cert');

    fs = require('fs');

    function ApiThrough() {
      var host, mongooose;
      mongooose = require('mongoose');
      host = process.env["MONGODB_PORT_27017_TCP_ADDR"] || 'localhost';
      console.log("Connecting to mongo at " + host);
      mongooose.connect("mongodb://" + host + ":27017/apis");
    }

    ApiThrough.prototype.start = function() {
      var proxy;
      proxy = new Proxy();
      proxy.use(this);
      proxy.onCertificateMissing = (function(_this) {
        return function(ctx, files, callback) {
          return _this.onCertificateMissing(ctx, files, callback);
        };
      })(this);
      proxy.listen({
        port: process.env['PROXY_PORT'] || 9081,
        sslCertCacheDir: './scripts/certs/http-mitm-proxy'
      });
      return this.proxy = proxy;
    };

    ApiThrough.prototype.onCertificateMissing = function(ctx, files, callback) {
      var sslCert;
      console.log('Looking for "%s" certificates', ctx.hostname);
      console.log('"%s" missing', ctx.files.keyFile);
      console.log('"%s" missing', ctx.files.certFile);
      sslCert = new SSLCert(ctx.hostname);
      return sslCert.create((function(_this) {
        return function() {
          return callback(null, {
            keyFileData: fs.readFileSync(ctx.files.keyFile),
            certFileData: fs.readFileSync(ctx.files.certFile)
          });
        };
      })(this));
    };

    ApiThrough.prototype.onError = function(ctx, err) {
      console.error('proxy error:', err);
      return console.error('proxy error stack:', err.stack);
    };

    ApiThrough.prototype.onRequest = function(ctx, callback) {
      var ApiExample, User, apiExample, responseAggregator, responseBody;
      ApiExample = require('./api_example');
      User = require('./user');
      ctx.onError((function(_this) {
        return function(ctx, err) {
          return _this.onError(ctx, err);
        };
      })(this));
      apiExample = new ApiExample();
      apiExample.populateFromRequest(ctx.clientToProxyRequest);
      apiExample.setFullUrl(ctx.isSSL, ctx.proxyToServerRequestOptions);
      responseBody = '';
      responseAggregator = new PassThrough();
      responseAggregator.on('finish', function() {
        apiExample.responseBody = responseBody;
        apiExample.stripResponseBody();
        return User.findOne({
          api_token: apiExample.apiToken
        }, function(err, user) {
          var apiExampleRaw;
          console.log('findOne', err, user);
          if (user) {
            apiExample.userId = user.id;
          }
          apiExampleRaw = apiExample.toObject();
          delete apiExampleRaw._id;
          return ApiExample.findOneAndUpdate({
            digest: apiExample.digest
          }, apiExampleRaw, {
            upsert: true
          }, function(error) {
            if (error != null) {
              return console.log("Failed to save due to error", error);
            }
          });
        });
      });
      ctx.addResponseFilter(responseAggregator);
      ctx.onRequestData(function(ctx, chunk, callback) {
        apiExample.requestBody += chunk.toString('utf8');
        return callback(null, chunk);
      });
      ctx.onResponse(function(ctx, callback) {
        apiExample.responseHeaders = ctx.serverToProxyResponse.headers;
        apiExample.statusCode = ctx.serverToProxyResponse.statusCode;
        ctx.serverToProxyResponse.on("finish", function() {
          return console.log("FINISH");
        });
        return callback();
      });
      ctx.onResponseData(function(ctx, chunk, callback) {
        responseBody += chunk.toString('utf8');
        return callback(null, chunk);
      });
      return callback();
    };

    return ApiThrough;

  })();

}).call(this);

\end{lstlisting}

\begin{lstlisting}[caption=lib/api\_example.js]

(function() {
  var ApiExample, ApiExamplesSchema, CUSTOM_HEADERS, VERSION_IN_HEADER, VERSION_IN_URL, crypto, inflector, mongoose, url, validator, _u;

  mongoose = require('mongoose');

  url = require('url');

  _u = require('underscore');

  crypto = require('crypto');

  validator = require('validator');

  inflector = require('i')();

  VERSION_IN_HEADER = /v(\d|\.\d)+/;

  VERSION_IN_URL = /\/(v\d[^\/]*)/;

  CUSTOM_HEADERS = {
    DESC_HEADER: "x-spy-rest-desc",
    VERSION_HEADER: "x-spy-rest-version",
    RESOURCE_HEADER: "x-spy-rest-resource",
    API_TOKEN_HEADER: "x-spy-rest-api-token",
    HOST_HEADER: 'x-spy-rest-host',
    NO_STRIP_HEADER: 'x-spy-rest-no-strip',
    ACTION_HEADER: 'x-spy-rest-action'
  };

  ApiExamplesSchema = new mongoose.Schema({
    description: {
      type: String
    },
    version: {
      type: String,
      "default": 'Default'
    },
    resource: {
      type: String
    },
    action: {
      type: String
    },
    url: {
      type: String
    },
    templatedURL: {
      type: String
    },
    query: {
      type: Object
    },
    host: {
      type: String
    },
    http_method: {
      type: String
    },
    requestBody: {
      type: String,
      "default": ''
    },
    requestHeaders: {
      type: Object
    },
    statusCode: {
      type: Number
    },
    responseHeaders: {
      type: Object
    },
    responseBody: {
      type: String,
      "default": ''
    },
    strippedResponseBody: {
      type: String
    },
    recordedAt: {
      type: Date,
      "default": function() {
        return new Date();
      }
    },
    fullURL: {
      type: String,
      "default": ''
    },
    digest: {
      type: String,
      "default": ''
    },
    requiresAuth: {
      type: Boolean,
      "default": false
    },
    apiToken: {
      type: String,
      "default": ''
    },
    userId: {
      type: mongoose.Schema.Types.ObjectId,
      "default": null
    },
    authType: {
      type: String
    }
  }, {
    collection: 'api_examples'
  });

  ApiExamplesSchema.index({
    host: 1,
    version: 1,
    resource: 1,
    action: 1
  });

  ApiExample = mongoose.model('ApiExample', ApiExamplesSchema);

  ApiExample.prototype.populateFromRequest = function(request) {
    this.host = request.headers[CUSTOM_HEADERS.HOST_HEADER] || request.headers.host;
    this.url = this.filteredUrl(request.url);
    this.http_method = request.method;
    this.requestHeaders = request.headers;
    this.description = request.headers[CUSTOM_HEADERS.DESC_HEADER];
    this.version = request.headers[CUSTOM_HEADERS.VERSION_HEADER] || this.guessedVersion() || 'Default';
    this.resource = request.headers[CUSTOM_HEADERS.RESOURCE_HEADER] || this.guessedResource();
    this.query = this.parsedUrl().query;
    this.apiToken = this.requestHeaders[CUSTOM_HEADERS.API_TOKEN_HEADER];
    this.filterAuthHeaders();
    this.action = this.computedAction();
    return this.digest = this.computeDigest();
  };

  ApiExample.prototype.stripResponseBody = function() {
    var StrippedObject, strippedObject;
    if (!this.shouldStrip()) {
      return;
    }
    StrippedObject = require('./stripped_object');
    strippedObject = new StrippedObject();
    try {
      return this.strippedResponseBody = JSON.stringify(strippedObject.strip(JSON.parse(this.responseBody)));
    } catch (_error) {
      return this.strippedResponseBody = this.responseBody;
    }
  };

  ApiExample.prototype.saveWithErrorLog = function() {
    return this.save(function(error) {
      if (error) {
        return console.error("Failed to save because of error", error);
      }
    });
  };

  ApiExample.prototype.guessedVersion = function() {
    return this.guessedVersionFromURL() || this.guessedVersionAcceptHeader();
  };

  ApiExample.prototype.guessedVersionFromURL = function() {
    if (this.url == null) {
      return null;
    }
    if (this.url.match(VERSION_IN_URL)) {
      return RegExp.$1;
    }
  };

  ApiExample.prototype.guessedVersionAcceptHeader = function() {
    var matches;
    if (this.requestHeaders['accept'] == null) {
      return null;
    }
    matches = this.requestHeaders['accept'].match(VERSION_IN_HEADER);
    if (matches != null) {
      return matches[0];
    }
  };

  ApiExample.prototype.parsedUrl = function() {
    return url.parse(this.url, true);
  };

  ApiExample.prototype.guessedResource = function() {
    var parts, possibleResource;
    parts = this.parsedUrl().pathname.split('/').reverse();
    possibleResource = _u.find(parts, function(part) {
      return part.match(/^[^\d].*$/);
    });
    if (possibleResource != null) {
      return possibleResource.split('.')[0];
    }
  };

  ApiExample.prototype.computedAction = function() {
    if (this.requestHeaders[CUSTOM_HEADERS.ACTION_HEADER] != null) {
      return "" + this.http_method + " " + this.requestHeaders[CUSTOM_HEADERS.ACTION_HEADER];
    } else {
      this.templatizeURL();
      return "" + this.http_method + " " + this.templatedURL;
    }
  };

  ApiExample.prototype.setFullUrl = function(isSSL, hostPort) {
    var host, port, scheme;
    scheme = isSSL ? "https" : "http";
    host = hostPort.host;
    port = '';
    if (hostPort.port) {
      if (scheme === "http" && hostPort.port !== 80) {
        port = ":" + hostPort.port;
      } else if (scheme === "https" && hostPort.port !== 443) {
        port = ":" + hostPort.port;
      }
    }
    return this.fullURL = "" + scheme + "://" + host + port + this.url;
  };

  ApiExample.prototype.filterAuthHeaders = function() {
    return this.requestHeaders = _u.reduce(this.requestHeaders, (function(_this) {
      return function(filteredHeaders, value, key) {
        if (key === 'authorization') {
          console.log("value");
          console.log(value);
          if (value.match(/Basic/)) {
            console.log(value);
            _this.authType = 'Basic';
            filteredHeaders[key] = 'Basic FILTERED';
          } else {
            filteredHeaders[key] = 'FILTERED';
          }
          _this.requiresAuth = true;
        } else {
          filteredHeaders[key] = value;
        }
        return filteredHeaders;
      };
    })(this), {});
  };

  ApiExample.prototype.filteredUrl = function(rawUrl) {
    var api_key_param, parasedRawUrl;
    if (!rawUrl) {
      return rawUrl;
    }
    parasedRawUrl = url.parse(rawUrl, true);
    api_key_param = _u.find(_u.keys(parasedRawUrl.query), function(param) {
      return param.toLowerCase() === 'api_key';
    });
    if (!api_key_param) {
      return rawUrl;
    }
    return rawUrl.replace("" + api_key_param + "=" + parasedRawUrl.query[api_key_param], "" + api_key_param + "=FILTERED");
  };

  ApiExample.prototype.computeDigest = function() {
    var hash, text;
    text = "__VERSION__" + this.version + "__RESOURCE__" + this.resource + "__URL__" + this.action + "__DESC__" + this.description;
    hash = crypto.createHash('sha');
    console.log("digest for text " + text);
    hash.update(text);
    return hash.digest('base64');
  };

  ApiExample.prototype.shouldStrip = function() {
    var strip_header_value;
    if (!this.requestHeaders) {
      return true;
    }
    console.log("HERE");
    strip_header_value = this.requestHeaders[CUSTOM_HEADERS.NO_STRIP_HEADER];
    return strip_header_value !== 'true';
  };

  ApiExample.prototype.templatizeURL = function() {
    var parts, path, templatedParts;
    path = this.parsedUrl().pathname;
    parts = path.split("/");
    templatedParts = _u.map(parts, function(part, index) {
      var previousPart;
      previousPart = index > 0 ? parts[index - 1] : '';
      if (validator.isNumeric(previousPart)) {
        return part;
      }
      if (validator.isNumeric(part)) {
        return "{:" + (inflector.singularize(previousPart)) + "-id}";
      } else if (validator.isUUID(part)) {
        return "{:" + (inflector.singularize(previousPart)) + "-guid}";
      } else {
        return part;
      }
    });
    return this.templatedURL = templatedParts.join('/');
  };

  module.exports = ApiExample;

}).call(this);
\end{lstlisting}