### 2017-11-08 symbols.tex
overview + questionnaire

%%%%%%%%%%%%%%%%%%%%%%%
%Sample List of Symbols
%%%%%%%%%%%%%%%%%%%%%%%

\begin{tabbing}
Symbol~~~~~\= \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \  \parbox{5in}{Definition}\\

\addsymbol \mbox{API}: {Application Programming Interface}
\addsymbol \mbox{HTTP}: {Hyper Text Transfer Protocol}
\addsymbol \mbox{JSON}: {JavaScript Object Notation}
\addsymbol \mbox{REST}: {Representational State Transfer, used interchangeably with RESTful}
\addsymbol \mbox{SOAP}: {Simple Object Access Protocol}
\addsymbol \mbox{WSDL}: {Web Service Definition Language}
\addsymbol \mbox{XML}: {Extensible Markup Language}
\addsymbol \mbox{REST API Developer}: {A software developer who implements REST API systems}
\addsymbol \mbox{REST API Client Developer}: {A software developer who implements systems using REST APIs}

% .
% .
% .
% ALWAYS KEEP THE FOLLOWING LINE
\end{tabbing}